---
title: "DataFest"
author: "Albert Sun and Ben Wallace"
date: "4/9/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r upload packages, results = "hide"}
library(dplyr)
library(ggplot2)
library(patchwork)
library(twitteR)
library(lubridate)
library(rvest)
library(sf)
library(tidycensus)
library(mapview)
library(broom)
library(tidyverse)       # Tidyverse for Tidy Data
library(readxl)
library(tigris)
library(sf)
library(viridis)
```

```{r read files, results = "hide"}
counties <- read.csv("datafiles/us-counties.csv")
states <- read.csv("datafiles/us-states.csv")
measures <- read.csv("datafiles/statedata.csv")
global_measures <- read.csv("datafiles/ACAPSGlobalResponse.csv")
covidapproval <- read.csv("datafiles/Covidapproval_Population.csv")
jails <- read.csv("datafiles/DataFest Covid in Jails - Sheet1.csv")
```

```{r topten}
toptencounties <- counties %>%
  filter(date == "2020-04-08") %>%
  arrange(desc(cases))
  
toptencounties <- toptencounties[1:10,]

toptenstates <- states %>%
  filter(date == "2020-04-08") %>%
  arrange(desc(cases))

toptenstates <- toptenstates[1:10,]
```

# NY State and County Date

```{r plot}
p1 <- ggplot(data = toptencounties, mapping = aes(x = reorder(county, cases), 
                                                  y = cases), 
             fill=if_else(County=="New York")) +
  geom_bar(stat = "identity") + 
  labs(title = "Top 10 Coronavirus counties cases", 
       y = "Cases", x = "County Name") + 
  coord_flip()

p2 <- ggplot(data = toptenstates, mapping = aes(x = reorder(state, cases), 
                                                y = cases)) +
  geom_bar(stat = "identity") + 
  labs(title = "Top 10 Coronavirus states cases", 
       y = "Cases", x = "State Name") + 
  coord_flip() 

p1 + p2
```

# Global Measures

```{r countries_most_social_distancing}
global_measures %>%
  filter(CATEGORY == "Social distancing") %>%
  count(COUNTRY) %>%
  arrange(-n)
```

First Social Distancing Measure From top-10 quickest and top-10 slowest Country 
to test responsiveness:

```{r convert_dates_to_ordered_lubridate}
first_measures <- global_measures %>%
  filter(CATEGORY == "Social distancing") %>%
  mutate(date = dmy(DATE_IMPLEMENTED)) %>%
  arrange(COUNTRY, date) %>%
  select(COUNTRY, date) %>%
  group_by(COUNTRY) %>%
  slice(1) %>%
  ungroup() %>%
  select(COUNTRY, date) %>%
  arrange(date) %>%
  slice(-11:-n()-10)
```

```{r top_bottom_10_reactions}
ggplot(first_measures, mapping = aes(x = reorder(COUNTRY, date), y = date)) +
  geom_point() + coord_flip() + 
  labs(y = "1st Social Distancing Measure Date", x = "Country",
       title = "1st Social Distancing Implemented") + 
  theme_minimal(base_size = 14)
```

# Jail Data

```{r Overview of Data}
jails %>%
  summarise(n(), n_distinct(State)) 
```

We have data from 334 counties from 27 states.

```{r}
topTenJails <- jails %>%
  arrange(Percentage.Change) %>%
  select(State, County, Percentage.Change) %>%
  slice(1:10)

ggplot(data = topTenJails, mapping = aes(x = reorder(County, Percentage.Change), 
                                         y = abs(Percentage.Change) * 100)) + 
  geom_bar(stat = "identity") +
  coord_flip() + 
  labs(x = "County", y = "Percent Decrease in Prisons", 
       title = "Top Ten Drops in Incarceration Rates in County Jails") 
```

```{r bunch_of_graphs}
ggplot(data = jails, aes(x = Percentage.Change)) +
  geom_histogram()

ggplot(data = jails, aes(x = County.Profile)) +
  geom_histogram(binwidth = .025)

ggplot(data = jails, aes(x = State.Profile)) +
geom_histogram(binwidth = 5)

jails$Jurisdiction.Profile = jails$Jurisdiction.Profile..R.B.M.

ggplot(data = jails, aes(x = Jurisdiction.Profile)) +
  geom_bar()

ggplot(data = jails, aes(x = Jail.Incarceration.Rate.Per.100K)) +
geom_histogram()

ggplot(data = jails, aes(x = COVID.19.Cases)) +
  geom_histogram()

jails %>%
  summarise(mean = mean(COVID.19.Cases), ## Also log transform this?
            median = median(COVID.19.Cases)
  )

ggplot(data = jails, aes(x = log(Deaths.Per.100K.Residents))) + ### Log transform variable?
  geom_histogram()

```

```{r Step-wise Strategy for Model Fitting}
lm_full <- lm(Percentage.Change ~ County.Profile + State.Profile + 
                Jurisdiction.Profile, data = jails)

lm <- step(lm_full, direction = "backward", trace=FALSE) 

tidy(lm) %>%
  select(term, estimate, p.value)
```
#^Stepwise function uses similar strategy to adjusted R squared to find the best
#fit model with constrinas in mind. Here, it finds that Jurisdiction.Profile is 
#the most important relationship
```{r graph Jurisdiction.Profile vs. Percentage.Change:}
ggplot(data = jails, aes(x = Jurisdiction.Profile, y = Percentage.Change)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal()
```

This graph doesn't look very legit...

```{r diagnostic_plots}
plot_diagnostic <- function(x, data) {
  
  plot_residuals <- augment(x) %>%
    ggplot(aes(x = .fitted, y = .resid)) +
    geom_point() + 
    geom_hline(yintercept = 0, color = "blue", lty = 2) +
    labs(x = expression(hat(y)), y = "Residual Value", 
         title = "Residuals Plot") 
  
  plot_normality <- augment(x) %>%
    ggplot(aes(x = .resid)) +
    geom_histogram(binwidth = .01, alpha = .5, color = "orange") + 
    labs(x = "Residual Value", y = "Count", title = "Normality Plot") 
  
  plot_independence <- augment(x) %>%
    ggplot(aes(x = 1:nrow(data), y = .resid)) + 
    geom_point() + 
    labs(x = "Index", y = "Residual Value", title = "Independence Plot") 
  
  plot_residuals / (plot_normality + plot_independence)
}

plot_diagnostic(lm, jails)
```
Conditions of Linearity:

Condition 1: We must verify that observations are independent. In this case, 
there doesn’t seem to be a pattern in the residual scatter plot, so the 
observations are independent.

Conditions 2 & 4: We must verify that the distribution of residuals is around 0 
and has constant variance. The variance plot shows that the distribution of 
residuals is centered around 0 and has constant variance throughout.

Condition 3: We must verify the normality of residuals. The residual histogram 
displays a normal distribution of residual values, so we have verified the 
normality of residuals.

```{r Lets try county.profile now}
lm(Percentage.Change ~ County.Profile, data = jails) %>%
  tidy()

ggplot(data = jails, aes(x = County.Profile, y = log(abs(Percentage.Change)))) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal()
```
Doesn't seem very legit either...